LAB 8 EXAMPLE
```{r}
X <- 1000 # number of samples
mu <- 5 # population mean
SD <- 2.5 # population standard deviation
y <- rnorm(n = X, mean = mu, sd = SD) # simulate the data
hist(y, col="lightblue", breaks=seq(-20,40,1)) # plot the data
mean(y) # sample mean
sd(y) # sample sd
Nmod <- lm(y ~ 1)
summary(Nmod)
length(y[which(y < 0)]) # y < 0 returns T of F
```

```{r}
n <- rpois(X, lambda = 5)
sort(unique(n))
table(n)
hist(n, col="lightblue", breaks=seq(-20,40,1))
length(n[which(n<0)])
mean(n)
var(n)
```

```{r}
par(mfrow=c(3,3))
means <- c(3,5,7,9,11,13,15,17,30)
for(i in 1:length(means)){
tmp.n <- rpois(500,means[i])
hist(tmp.n, breaks=seq(0,50,1), border="lightblue", col="lightblue",
xlim=c(0,50), ylim=c(0,120), main=paste("lambda = ",means[i]))
}
```

```{r}
Pmod <- lm(n ~ 1)
summary(Pmod)
par(mfrow=c(2,2))
#residuals from a 'normal' model with 'normal' data
clr <- adjustcolor(4,0.3)
plot(resid(Nmod), main="Normal residuals", pch=16, col=clr)
abline(h=0,lwd=2)
qqnorm(resid(Nmod), main="Normal residuals", pch=16, col=clr)
qqline(resid(Nmod), lwd=2)
#residuals from a 'normal' model with 'non-normal' data
clr <- adjustcolor(3,0.3)
plot(resid(Pmod), main="Poisson residuals", pch=16, col=clr)
abline(h=0,lwd=2)
qqnorm(resid(Pmod),main="Poisson residuals", pch=16, col=clr)
qqline(resid(Pmod), lwd=2)
```

```{r}
library(DHARMa)
simulated.resids.Pmod <- simulateResiduals(fittedModel = Pmod, n = 100)
plot(simulated.resids.Pmod)
```

"With ecological data, and as we talked about in class, we often violate the assumption that
variance equals the mean, in reality, the variance often exceeds the mean. Such data are
said to be overdispersed. Overdispersion often happens when data are clumped or otherwise
show lack of independence (think of data points taken close in space or time) and primarily
acts to increase standard errors."

NEGATIVE BINOMIAL DISTRIBUTION

```{r}
library(MASS)
nb <- rnegbin(X, mu=5, theta=100)
cc <- rpois(X, 5)
sort(unique(nb))
table(nb)
length(nb[which(nb<0)])
mean(nb)
var(nb)
#compare Poisson and Negative binomial
hist(nb, col="lightblue", breaks=seq(-10,20,1),
main="Negative Binomial, mu=5, k=100")
hist(cc, col=adjustcolor("red",0.3), breaks=seq(-10,20,1),
main="Poisson, lambda=5",add=T)
```

```{r}
par(mfrow=c(1,2))
theta <- c(100,10,1)
for(i in 1:3){
sim.nb <- rnegbin(X, mu=5, theta=theta[i])
sim.p <- rpois(X, 5)
mean(sim.nb)
var(sim.nb)
#compare Pois and Negbin
hist(sim.nb, col="lightblue", breaks=seq(-10,40,1),
main=paste("Negative Binomial, mu=5, k=",theta[i],sep=""))
hist(sim.p, col=adjustcolor("red",0.3), breaks=seq(-10,40,1),
main="Poisson, lambda=5",add=T)
}
```

```{r}
nb.1 <- rnegbin(X, mu=5, theta=1)
NBmod <- lm(nb.1 ~ 1)
summary(NBmod)
par(mfrow=c(2,2))
#residuals from a 'correct' model
clr <- adjustcolor(4,0.3)
plot(resid(Nmod), main="Normal residuals", pch=16, col=clr)
abline(h=0,lwd=2)
qqnorm(resid(Nmod), main="Normal residuals", pch=16, col=clr)
qqline(resid(Nmod))
#residuals from a 'mis-specified' model
clr <- adjustcolor(3,0.3)
plot(resid(NBmod), main="Negative binomial residuals", pch=16, col=clr)
abline(h=0,lwd=2)
qqnorm(resid(NBmod),main="Negative binomial residuals", pch=16, col=clr)
qqline(resid(NBmod))
```

BINOMIAL DISTRIBUTION

```{r}
z <- rbinom(X, size=1, prob=0.5)
sort(unique(z))
table(z)
length(z[which(z==0)]) # number of 0's
length(z[which(z==1)]) # number of 1's
mean(z) # proportion of ones (~p)
par(mfrow=c(1,1))
hist(z, breaks=seq(0,1,0.1), col="lightblue")
```

```{r}
z2 <- rbinom(X,size=1,prob=0.8)
sort(unique(z2))
table(z2)
hist(z2,breaks=seq(0,1,0.1),col="lightblue")
```

```{r}
z3 <- rbinom(X, size=4, prob=0.5)
sort(unique(z3))
table(z3)
hist(z3,breaks=seq(0,4,0.25),col="lightblue")
```

```{r}
par(mfrow=c(3,3))
trials <- c(1,3,6)
p <- c(0.3,0.5,0.8)
for(i in 1:length(trials)){
for(j in 1:length(p)){
tmp.z <- rbinom(500, trials[i], p[j])
hist(tmp.z, breaks=seq(-0.5,6.5,1), col="lightblue", xlim=c(-1,7),right=F,
ylim=c(0,400), main=paste("n = ",trials[i],"p = ",p[j]))
}
}
```

```{r}
Bmod <- lm(z~1)
summary(Bmod)
par(mfrow=c(1,1))
clr <- adjustcolor(4,0.3)
qqnorm(resid(Bmod), pch=16, col=clr)
qqline(resid(Bmod), lwd=2)
```

```{r}
samples <- 100
X <- runif(samples,0,100)
b0 <- 0
b1 <- 0.25
SD <- 1.5
par(mfrow=c(1,3))
#simulate response variable (rv)
rv.n <- rnorm(n=samples, mean= b0+b1*X, sd=SD)
rv.p <- rpois(n=samples, lambda= b0+b1*X)
rv.nb <- rnegbin(n=samples, mu= b0+b1*X, theta=1)
#Plot the data
plot(X, rv.n, xlim=c(-10,100), ylim=c(-10,50), pch=21, bg="lightblue",
cex=1.5, main="Normal errors")
abline(lm(rv.n~X), lwd=2, col="blue")
abline(h=0, lwd=2)
plot(X, rv.p, xlim=c(-10,100),ylim=c(-10,50), pch=21, bg="lightgreen",
cex=1.5, main="Poisson errors")
abline(lm(rv.p~X), lwd=2, col="darkgreen")
abline(h=0, lwd=2)
plot(X, rv.nb, xlim=c(-10,100), ylim=c(-10,50), pch=21, bg="yellow",
cex=1.5, main="Negative binomial errors")
abline(lm(rv.nb~X), lwd=2, col="orange")
abline(h=0, lwd=2)
```

```{r}
par(mfrow=c(3,2))
# fit models
Nmod <- lm(rv.n~X)
Pmod <- lm(rv.p~X)
NBmod <- lm(rv.nb~X)
# check summaries
summary(Nmod)
summary(Pmod)
summary(NBmod)
# diagnostic plots
plot(Nmod, 1:2, pch=21, cex=1.5, bg="lightblue")
plot(Pmod,1:2, pch=21, cex=1.5, bg="lightgreen")
plot(NBmod,1:2, cex=1.5,pch=21, bg="yellow")
```

```{r}
log.rv.nb <- log(rv.nb + 0.1)
logNBmod <- lm(log.rv.nb ~ X)
predlogNB <- exp(predict(logNBmod))
par(mfrow=c(1,1))
hist(predlogNB, col="lightblue", breaks=seq(-20,150,2), ylim=c(0,40),main="")
hist(rv.nb+0.1, col=adjustcolor(2,0.3), breaks=seq(-20,150,2), add=T)
legend("topright", c("Predicted","Actual"), pch=15, pt.cex = 2, bty="n",
col=c("lightblue",adjustcolor(2,0.3)))
```