Making Datq
```{r, include = TRUE}
set.seed(123)
covariate <- runif(100,0,50)
group <- rep(c("A", "B"), each=50)
intA <- 100
intB <- 150
slopeA <- 6
sigma <- 60
pars <- c(intA, intB, slopeA)
dm <- model.matrix(~group+covariate)
response <- rnorm(100, dm %*% pars, sigma)
sim.dat <- data.frame(response, group, covariate)
plot(response ~ covariate, col = c(1,2))
```


What is the slope for both groups?
-Looks positive
-Difference in intercept

```{r, include = TRUE}
Manc <- lm(response ~ group + covariate, data = sim.dat)
summary(Manc)
```
GroupA estimate: intercept for group A
GroupB esimate: difference in intercept between group A and B
covariate estimate: relationship between response and explanatory variable (slope)
Compared to created data, model didn't do too bad

```{r, include = TRUE}
plot(response~covariate, col = as.factor(group))
```

```{r, include = TRUE}
set.seed(123)
covariate <- runif(100,0,50)
group <- rep(c("A", "B"), each=50)
intA <- 100
intB <- 150
slopeA <- 6
slopeB <- -5
sigma <- 40
pars <- c(intA, intB, slopeA, slopeB)
dm <- model.matrix(~group*covariate)
response <- rnorm(100, dm %*% pars, sigma)
sim.dat <- data.frame(response, group, covariate)
plot(response ~ covariate, col = c(1,2))
```

slope for group B should be -1

Fit our model

```{r, include = TRUE}
Mint <- lm(response ~ group * covariate, data = sim.dat)
summary(Mint)
est <- coef(Mint)
plot(response ~ covariate, col = as.factor(group))
abline(100,6,lwd = 1,lty=2,col=2)
abline(100+150,1,lwd = 1,lty=2,col=4)

abline(est[1],est[3],lwd=2, col=2)
abline(est[1] + est[2], est[3] + est[4],lwd=2, col=4)

```

Lines represent what we told R lines should be vs what R calculated for the lines

Intercept estimate -> intercept of group A
groupB estiamte -> estimate difference between group A and B
covariate estimate -> slope between group A and response variable
groupB:covariate -> slope difference between group A and B

```{r, include = TRUE}

```


