---
editor_options: 
  markdown: 
    wrap: 72
---

Using horseshoe crab data that is presence and absence data.

```{r}
crabs <- read.table("crabs.txt", h=T)
crabs$AirTemp2 <- crabs$AirTemp^2
crabs$Binary <- ifelse(crabs$Crabs > 0, 1, 0)
```

ifelse is an if else statement, where (statement, return value for if,
return value for else)

1.  State the question/hypothesis

Can variation the probability of observing a crab at a site be explained
by: -Site: does crab probability vary across sites - intercept
differences -Temp: does crab probability of being present vary with
temp - slope differences

Data type: binary error distribution: binomial Binomial ANCOVA model

2.  Data exploration

```{r}
head(crabs)
require(ggplot2)
require(GGally)
ggpairs(crabs[,c("Site", "AirTemp", "Binary")])
```

Top left to bottom right graphs (thru line) represent number of
observations for that variable (binary is probability of seeing a crab)

```{r}
ggplot(crabs, aes(x = AirTemp, y = Binary, colour = Site)) +
  geom_jitter(size = 2, alpha = 0.4, position = position_jitter(height = 0.02)) +
  stat_smooth(method = "loess", color = "blue") +
  theme_light()
```

We might consider adding an x\^2 component because we see sort of a
peak, but not 100%. We just may not have captured the high temperature
range.

3.  Describe the model

Linear predictor: n~i~ = B~o~ + B~site~X~sitei~ + B~air~X~airi~ logit
link binomial error distribution

We can add a quadratic to see if it's meaningful.

Linear predictor: n~i~ = B~o~ + B~site~X~sitei~ + B~air~X~airi~ +
B~air2~X~air2i~

Assumptions: Bionomial dtributions no error

Check for collinearity

```{r}
require(car)
m1 <- glm(Crabs ~ Site + AirTemp, data = crabs, family = poisson())
vif(m1)
```

Below 2, which is good.

4)  Fit the model

```{r}
require(AICcmodavg)
fitList <- list(
  "null"     = glm(Binary~1,                      data = crabs, family = binomial), #Null
  "Site"     = glm(Binary~Site,                   data = crabs, family = binomial), #ANOVA
  "Temp"     = glm(Binary~AirTemp,                data = crabs, family = binomial), #regression
  "Temp2"    = glm(Binary~AirTemp+AirTemp2,       data = crabs, family = binomial), #mult. reg.
  "SiteTemp" = glm(Binary~Site+AirTemp,           data = crabs, family = binomial), #ANCOVA
  "SiteTemp2"= glm(Binary~Site+AirTemp+AirTemp2,  data = crabs, family = binomial)  #ANCOVA
)

aictab(fitList, modnames = names(fitList))
```

Both SiteTemp and Temp have very close results, typically you should
choose the one with lower K (simplier model); however, we will use
SiteTemp to understand site differences as well.

5)  Evaluate the output

```{r}
m.best <- fitList[["SiteTemp"]]
print(summary(m.best),show.residuals=TRUE)
```

Estimate Intercept is probability of presence of Crabs at Erika's when
airtemp = 0. Next coefficients are constrasts (they are the intercepts).
The Airtemp estimate is the slope for probability of crab based on one
increment of airtemp.

The Chi squared value was 0.61.

Differences across sites are not too large, but there seems to be a
relationship between airtemp and probability of crab presence.

Let's look at temp since there are no signficant differences for sites

```{r}
m.best <- fitList[["Temp"]]
print(summary(m.best),show.residuals=TRUE)
```

Different p value and different chi squared goodness of fit value
(0.38). Chi squared goodness of fit may help describe how well you model
fits, but not an exact 1 to 1 description.

We can also use AUC (area under the curve). Another nice way of
determining fit for presence or absence data. AUC is area under the ROC
curve. ROC curve plots the true positive rate vs the flase positive
rate. Ranges from 0 to 1, 1 being perfect classifier (predicted
perfectly), 0 being wrong (poorly predicted)

```{r}
library(pROC)
myroc <- roc(crabs$Binary, predict(m.best, crabs, type= "response"))
plot(myroc)

ggroc(myroc, legacy.axes = TRUE) + 
  labs(x = "False-positive rates", y = "True-Positive Rate", title = "Simulated ROC curve for Temp model") + annotate("text", x = .5, y = .5, label = paste0("AUC: ", round(auc(myroc), digits = 2)))
```

Can use this to compare between different modles (close to 1 the
better), null model should be 0.5.

```{r}
pred.df <- data.frame(AirTemp = seq(min(crabs$AirTemp), max(crabs$AirTemp), length=300))

#predictions on the RESPONSE scale!
pred <- predict(m.best, newdata = pred.df, se.fit = TRUE, type = "response")
pred.df$fit <- pred$fit
pred.df$se.fit <- pred$se.fit

# make CI (by hand since package not working well)
pred.df$resp.lo <- pred.df$fit - 1.96*pred.df$se.fit
pred.df$resp.hi <- pred.df$fit + 1.96*pred.df$se.fit

require(ggplot2)

ggplot(pred.df, aes(x = AirTemp, y = fit)) + 
  geom_line(color = "blue") + 
  geom_ribbon(aes(ymin=resp.lo, ymax=resp.hi), fill = "blue", alpha=0.1) +
  geom_jitter(data=crabs, aes(x = AirTemp, y = Binary), size = 2, alpha = 0.4, position = position_jitter(height = 0.02)) +
  ylab("Probability of presence") +
  xlab("Temp (C)") +
  theme_classic()
```

6)  Interpret the Results

Does not differ between sites, increases probability of crabs with
temperature. No support for quadratic effect, which is not biologically
reasonable (temperature cannot infinitely rise and number of crabs too)
