---
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = TRUE}
snakes <- read.table("indigos.txt", h=T)
snakes$log.HR = log(snakes$hr.size)
```

Question: Does habitat composition at the home range scale explain
variation in home range size?

```{r, include = TRUE}
mG <- glm(log.HR ~ urban1.50 + upland1.50 + wetland1.50, data = snakes)
model.matrix(~ urban1.50 + upland1.50 + wetland1.50, data = snakes)
summary(mG)
mTop <- glm(log.HR ~ urban1.50 + upland1.50, data = snakes )
```

Math model:

Y~i~ = B~o~ + B~1~X~urbi~ + B~2~X~upi~ + B~3~X~wet~ + e~i~

Looking at the summary: The residuals are "ok" and may seem normal. When
we have no urban, wetland, or upland -\> territory size should be 4.6
For every one unit urban increase, there is a 1.8 decrease unit in
territory size "going 0 to all upland habitat" you increase by 0.7171
territory size "going 0 to all wetland habitat" you increase by 0.6711
territory size

Create models with all 3, some with only 1, and some with
combinations... The first 3 models were all similar, these ones were...

Urban + Upland Urban Urban + Upland + Wetlands

```{r, include = TRUE}
summary(mTop)
```

Urban now shows a decrease of 2 per 0-100 habitat. 

Is per unit 1% or 0 to 100%?

Confidence interval -> observation count
