---
editor_options: 
  markdown: 
    wrap: 72
---

```{r include = TRUE}
getwd()
mander <- read.table(file = "salamander.txt", h=T)
mander$tail <- mander$TL-mander$SVL
```

1.  State the Question

Does tail length scale predictable with SVL?

H~o~: There is no relationship between tail length and SVL -response:
tail length -Explanatory: SVL

2.  Data exploration

```{r, include = TRUE}
plot(tail~SVL, data = mander)
```

What kind of parameters (B~o~ and B~1~) do you expect?

Makes sense for B~o~ to be 0 or close to 0. B~1~ looks close to one.

3.  Describe the model

In words: Is there a significant relationship between TL and SVL?

In math: Y~i~ + B~1~X~1i~ + e~i~

4.  Fit the model

```{r include = TRUE}
mALLo <- lm(tail ~ SVL, data = mander)
summary(mALLo)
```

Place in Equation: Y = 4.99 + 0.86x, relationship between x and y.

5.  Evaluate the Model

Model validations - check assumptions! -Residuals normal -Constant
variance -Observations are independent -Predictors measured without
error (fixed X)

```{r, include = TRUE}
par(mfrow = c(2,2))
plot(mALLo)
```

Residuals vs Fitted are not a straight line and therefore not the best
fit for the model. QQ plot is alright , some spots at the end veer off
the straight line. Scale location plot is OK, but some stragglers are a
bit concerning. Residuals vs leverage indicates that potential outlines
do not have high leverage and don't have a strong influence on
parameters. For every unit increase of SVL the tail will increase 0.89 mm.

Model selection: AIC table

6. Interpret the results

-Use predict to show relationship AND CI

```{r, include = TRUE}
new.df <- data.frame(SVL = mander$SVL)
pred <- predict(mALLo, newdata=new.df, interval = "confidence")

plot(mander$tail~mander$SVL)
lines(pred[,1] ~ new.df$SVL, lwd = 2)
lines(pred[,2] ~ new.df$SVL, lty = 2, col = "red")
lines(pred[,3] ~ new.df$SVL, lty = 2, col = "red")
```

How are the confidence intervals calculated? Why aren't they the same all the wall along the model line?

The more data we have around the model, the more confident we are in that model mean. 

Does the intercept makes biological sense?

Not particularly... Ways to standardize data:

Centering
-Centering data around 0 by subtracting the mean.
-Units remain the same

Normalization
-Subtracting the mean and dividing by the sample SD
-One unit is now a standard deviation (a bit harder to interpret)

By creating a new model with the centered or normalized data then the intercept will change to make it more biologically relevant. 
When using normal changed data -> slope is the change in tail length for every 1 SD change in SVL. 

Multiple Linear Regression
Y~i~ = B~o~ + B~1~X~1i~ + B~2~X~2i~ + e~i~

