```{r}
require(MASS)
set.seed(123)

n <- 100
X <- sort(runif(n, 0, 4)) #continuous predictor variable
b0 <- -1 #intercept
b1 <- 1 #slope: change in y on LP scale
lin.pred <- b0 + b1*X #linear predictor
eY <- exp(lin.pred) #exp count: response scale

c.P <- rpois(n = n, lambda = eY) #poisson distribution
c.NB <- rnegbin(n=n, mu = eY, theta = 2.5) #negative binomial distribution

m.pois <- glm(c.P ~ X, family = poisson)
summary(m.pois)$coefficients
```

For every unit of X increase, we increase log y (1.015) because it's still on a linear predictor scale. Changing it to a response scale the graph would look exponential.

model mean of y = e\^(B~o~ + B~1~X~1~)

Every time we increase by a unit of X, we increase the predict value of y by B~1~

```{r}
summary(m.pois)
```

Fit a model that assumes poisson distribution with data that was created for poisson. 

How do we know how adequate our model is? -> deviances

```{r}
dev <- summary(m.pois)$deviance
d.o.f <- summary(m.pois)$df.residual

#Chi-squared test (>0.05 indicates adequate fit)
1-pchisq(dev, d.o.f)
```
Looking at the value, it does pretty well. 

```{r}
m.negbin <- glm(c.NB ~ X, family = poisson)

dev.nb <- summary(m.negbin)$deviance
dof.nb <- summary(m.negbin)$df.residual

1-pchisq(dev.nb, dof.nb)
```

Not good value... That's what we get for trying to fit a poisson distribution to a negative binomial. 

```{r}
require(DHARMa)
m.pois.simRsid <- simulateResiduals(fittedModel = m.pois, plot = F)
plot(m.pois.simRsid)
```

Use DHARMa package because trying to compare GLM with link function looks at non transformed scale, but the package allows you to look at residuals that are transformed. 

The first one is a QQplot, observered vs expected. Looking for a linear line. 

Residuals vs predicted plot: a tiny bit of convergence but not terrible, hope is that each line (quartile lines) follow the dotted lines. 

When plotting the residuals of the poisson GLM with NB data you see that "S" pattern in the qqplot and the residuals graph is not flat at all. 

Over dispersed data (bad), residuals do not match poisson distribution, too much variation

```{r}
summary(m.pois)
```

When residual deviance is close to degreees of freedom it is close to one which is good! Deviance/degrees of freedom should equal 1 (P)

```{r}
summary(m.negbin)
```

Bad P value

We can calculate the signficant of the overdispersion using dispersiontest() from libraer (AER)

```{r}
require(AER)
dispersiontest(m.pois)
dispersiontest(m.negbin)
```

Closer to one, the better the value. Value is signficantly overdispersion...

```{r}
mNB <- glm.nb(c.NB ~ X)
dev <- summary(mNB)$deviance
dof <- summary(mNB)$df.residual
1-pchisq(dev, dof)
summary(mNB)$deviance / summary(mNB)$df.residual
```

Chisquared - how well your model fits, dispersion parameter, closer to one which is good, model fits residuals, looking good!

```{r}
aic.df <- data.frame(Model = c("Pois", "negBin"),
  AIC = c(AIC(m.pois), AIC(mNB)),
  dAIC = c(AIC(m.pois), AIC(mNB)) - min(c(AIC(m.pois), AIC(mNB))))
(aic.df <- aic.df[order(aic.df$AIC),])
```

```{r}

```