```{r, include = TRUE}
crab <- read.table("crabs.txt", h=T)
```

1. State the Question

Can variation in relative crab abundance be explained by: Site and Temperature

Does crab abundance vary across sites AND does crab abundance vary with temperatures?

Sites- > Differences in intercepts
Temperature -> Difference in slopes

Data type: count
Error Distribution: Poisson
Therefore: Poisson ANCOVA (categorical and continous variable)

2. Data exploration

```{r, include = TRUE}
install.packages(GGalley)
library(GGalley)
ggpairs(crab)
```

Found potential collinearity between air and water temperatures... Create global models and use VIF to see which variable we should keep.

```{r, include = TRUE}
library(car)
m1 <- glm(Crabs ~ Site + AirTemp + WaterTemp, data = crab, family = poisson)
vif(m1)
```

Air temp is slightly better than water temp and is easier to measure, so choose air temp in this case. Rule of thumb, GVIF < 2 is ideal. 

```{r, include = TRUE}
library(car)
m1 <- glm(Crabs ~ Site + AirTemp, data = crab, family = poisson)
vif(m1)
```

Now all values are below two and we are happy.

```{r, include = TRUE}
hist(crab$Crabs)
boxplot(crab$Crabs ~ crab$Site)
plot(crab$Crabs ~ crab$AirTemp)
```

Some variation by site, and there seems to be some window that allows crabs to thrive in (some optimum air temperature)

At many sites we don't see any crabs, a lot of zeros at our sites, but we also have a pretty long tail. We might expect a quadratic relationship!

3) Describe the model

Linear predictor: n~i~ + B~o~ + B~site~X~sitei~ + B~air~X~airi~
B~o~ is the intercept at the reference site when air temperature is 0
B~site~ are the 6 contrasts between each of the sites (when air temp = 0 )
B~air~ is the change in n with a unit increase in air temp (slope)



