---
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = TRUE}
crab <- read.table("crabs.txt", h=T)
```

1.  State the Question

Can variation in relative crab abundance be explained by: Site and
Temperature

Does crab abundance vary across sites AND does crab abundance vary with
temperatures?

Sites- \> Differences in intercepts Temperature -\> Difference in slopes

Data type: count Error Distribution: Poisson Therefore: Poisson ANCOVA
(categorical and continuous variable)

2.  Data exploration

```{r, include = TRUE}
install.packages(GGalley)
library(GGalley)
ggpairs(crab)
```

Found potential collinearity between air and water temperatures...
Create global models and use VIF to see which variable we should keep.

```{r, include = TRUE}
library(car)
m1 <- glm(Crabs ~ Site + AirTemp + WaterTemp, data = crab, family = poisson)
vif(m1)
```

Air temp is slightly better than water temp and is easier to measure, so
choose air temp in this case. Rule of thumb, GVIF \< 2 is ideal.

```{r, include = TRUE}
library(car)
m1 <- glm(Crabs ~ Site + AirTemp, data = crab, family = poisson)
vif(m1)
```

Now all values are below two and we are happy.

```{r, include = TRUE}
hist(crab$Crabs)
boxplot(crab$Crabs ~ crab$Site)
plot(crab$Crabs ~ crab$AirTemp)
```

Some variation by site, and there seems to be some window that allows
crabs to thrive in (some optimum air temperature)

At many sites we don't see any crabs, a lot of zeros at our sites, but
we also have a pretty long tail. We might expect a quadratic
relationship!

3)  Describe the model

Linear predictor: n~i~ + B~o~ + B~site~X~sitei~ + B~air~X~airi~ B~o~ is
the intercept at the reference site when air temperature is 0 B~site~
are the 6 contrasts between each of the sites (when air temp = 0 )
B~air~ is the change in n with a unit increase in air temp (slope)

Error/residuals distributions: e~i~ \~ Pois(lambda)

when lambda = u, this means our variance is equal to our mean (the mean being our modeled mean)

link function: n = g(u) = log(u)

Linear predictor: n~i~ + B~o~ + B~site~X~sitei~ + B~air~X~airi~ + B~air2~X~air2i~

```{r, include = TRUE}
crab$AirTemp2 <- (crab$AirTemp^2)
```

What are our assumptions?
-Residuals have a Poisson Distribution
-Observations are independent
-Measured without error

Fit the model
```{r, include = TRUE}
fitList <- list(
  m0 <- glm(Crabs~1, data = crab, family = poisson),
  m1 <- glm(Crabs~Site, data = crab, family = poisson),
  m2 <- glm(Crabs~AirTemp, data = crab, family = poisson),
  m3 <- glm(Crabs~AirTemp+AirTemp2, data = crab, family = poisson),
  m4 <- glm(Crabs~Site+AirTemp, data = crab, family = poisson),
  m5 <- glm(Crabs~Site+AirTemp+AirTemp2, data = crab, family = poisson)
)
```

5) Evaluate the model output

```{r, include = TRUE}
library(AICcmodavg)
aictab(fitList, modnames = paste0("m",0:5))
```