---
editor_options: 
  markdown: 
    wrap: 72
---

Poisson distribution and Negative Binomial example

```{r, include = TRUE}
crab <- read.table("crabs.txt", h=T)
```

1.  State the Question

Can variation in relative crab abundance be explained by: Site and
Temperature

Does crab abundance vary across sites AND does crab abundance vary with
temperatures?

Sites- \> Differences in intercepts Temperature -\> Difference in slopes

Data type: count Error Distribution: Poisson Therefore: Poisson ANCOVA
(categorical and continuous variable)

2.  Data exploration

```{r, include = TRUE}
install.packages(GGalley)
library(GGally)
ggpairs(crab)
```

Found potential collinearity between air and water temperatures...
Create global models and use VIF to see which variable we should keep.

```{r, include = TRUE}
library(car)
m1 <- glm(Crabs ~ Site + AirTemp + WaterTemp, data = crab, family = poisson)
vif(m1)
```

Air temp is slightly better than water temp and is easier to measure, so
choose air temp in this case. Rule of thumb, GVIF \< 2 is ideal.

```{r, include = TRUE}
library(car)
m1 <- glm(Crabs ~ Site + AirTemp, data = crab, family = poisson)
vif(m1)
```

Now all values are below two and we are happy.

```{r, include = TRUE}
hist(crab$Crabs)
boxplot(crab$Crabs ~ crab$Site)
plot(crab$Crabs ~ crab$AirTemp)
```

Some variation by site, and there seems to be some window that allows
crabs to thrive in (some optimum air temperature)

At many sites we don't see any crabs, a lot of zeros at our sites, but
we also have a pretty long tail. We might expect a quadratic
relationship!

3)  Describe the model

Linear predictor: n~i~ + B~o~ + B~site~X~sitei~ + B~air~X~airi~ B~o~ is
the intercept at the reference site when air temperature is 0 B~site~
are the 6 contrasts between each of the sites (when air temp = 0 )
B~air~ is the change in n with a unit increase in air temp (slope)

Error/residuals distributions: e~i~ \~ Pois(lambda)

when lambda = u, this means our variance is equal to our mean (the mean being our modeled mean)

link function: n = g(u) = log(u)

Linear predictor: n~i~ + B~o~ + B~site~X~sitei~ + B~air~X~airi~ + B~air2~X~air2i~

```{r, include = TRUE}
crab$AirTemp2 <- (crab$AirTemp^2)
```

What are our assumptions?
-Residuals have a Poisson Distribution
-Observations are independent
-Measured without error

Fit the model
```{r, include = TRUE}
fitList <- list(
  m0 <- glm(Crabs~1, data = crab, family = poisson),
  m1 <- glm(Crabs~Site, data = crab, family = poisson),
  m2 <- glm(Crabs~AirTemp, data = crab, family = poisson),
  m3 <- glm(Crabs~AirTemp+AirTemp2, data = crab, family = poisson),
  m4 <- glm(Crabs~Site+AirTemp, data = crab, family = poisson),
  m5 <- glm(Crabs~Site+AirTemp+AirTemp2, data = crab, family = poisson)
)
```

We need Airtemp with Airtemp2 always because y = x^2 + x is the shape we want not just y = x^2

5) Evaluate the model output

```{r, include = TRUE}
library(AICcmodavg)
aictab(fitList, modnames = paste0("m",0:5))

summary(m5)
```
We choose model 5 because it has the lowest AICc value. Site and airtemp is important when it comes to crab count. 

Looking at the summary table, the residuals are not too great. It would seem that the residuals are not very normally distributed. Airtemp is positively correlated, but Airtemp2 is negatively correlated. Originates from -X^2 + x = y (graph on desmos and formula works out). Each site coefficient is a contrast between the intercept of their own and the reference site (intercept being estimate of crab count when air temp = 0). 

Interpret the results

-Use models to predict differences in expected abundance among sites
-> Fix temperature at some value...
mean?
o?
value of management interest?
Others?

```{r}
require(DHARMa)
resids2 <- simulateResiduals(m5)
plot(resids2)
```

Bad plot! S curve is bad and lines on the right plot are curved. 

```{r, include = TRUE}
AirTemps = seq(min(crab$AirTemp), max(crab$AirTemp), length=300)
Sites <- levels(as.factor(crab$Site))

pred.df2 <- expand.grid(Site = Sites, AirTemp = AirTemps)
pred.df2$AirTemp2 <- pred.df2$AirTemp^2
pred2 <- predict(m5, newdata = pred.df2, se.fit = TRUE, type = "response")
pred.df2$fit <- pred2$fit
pred.df2$se.fit <- pred2$se.fit

# make CI (by hand since package not working well)
pred.df2$resp.lo <- pred.df2$fit - 1.96*pred.df2$se.fit
pred.df2$resp.hi <- pred.df2$fit + 1.96*pred.df2$se.fit

ggplot(pred.df2, aes(x = AirTemp, y = fit, color = Site, fill = Site)) + 
  geom_line() + 
  geom_ribbon(aes(ymin=resp.lo, ymax=resp.hi), alpha=0.1, colour = NA) +
  geom_point(data=crab, aes(x = AirTemp, y = Crabs), size = 2, alpha = 0.4) +
  ylab("Probability of presence") +
  xlab("Temp (C)") +
  theme_classic()
```

Why do we fix values?

```{r}
require(MASS)
m.null <- glm.nb(Crabs ~ 1, data = crab)
m.nb1 <- glm.nb(Crabs ~ Site, data = crab)
m.nb2 <- glm.nb(Crabs ~ Site + AirTemp, data = crab)
m.nb3 <- glm.nb(Crabs ~ Site + AirTemp + AirTemp2, data = crab)
m.nb4 <- glm.nb(Crabs ~ AirTemp, data = crab)
m.nb5 <- glm.nb(Crabs ~ AirTemp + AirTemp2, data = crab)

m.list <- list(
  "null" = m.null,
  "nb1" = m.nb1,
  "nb2" = m.nb2,
  "nb3" = m.nb3,
  "nb4" = m.nb4,
  "nb5" = m.nb5
)

aictab(m.list)

summary(m.list$nb3)
```

Cannot use regular QQ plot anymore for negative binomial, need to use a different diagonistc plot. Looking for NOT an S shape. Lines on the right should be straight (accounting for 0). If we see curves in the line -> over dispersion (typically too many 0s). A tiny bit of patterning in residual vs predicted, but too much.

```{r}
require(DHARMa)
resids2 <- simulateResiduals(m.nb3)
plot(resids2)
```

```{r, include = TRUE}
AirTemps = seq(min(crab$AirTemp), max(crab$AirTemp), length=300)
Sites <- levels(as.factor(crab$Site))

pred.df3 <- expand.grid(Site = Sites, AirTemp = AirTemps)
pred.df3$AirTemp2 <- pred.df2$AirTemp^2
pred3 <- predict(m.nb3, newdata = pred.df3, se.fit = TRUE, type = "response")
pred.df3$fit <- pred3$fit
pred.df3$se.fit <- pred3$se.fit

# make CI (by hand since package not working well)
pred.df3$resp.lo <- pred.df3$fit - 1.96*pred.df3$se.fit
pred.df3$resp.hi <- pred.df3$fit + 1.96*pred.df3$se.fit

ggplot(pred.df3, aes(x = AirTemp, y = fit, color = Site, fill = Site)) + 
  geom_line() + 
  geom_ribbon(aes(ymin=resp.lo, ymax=resp.hi), alpha=0.1, colour = NA) +
  geom_point(data=crab, aes(x = AirTemp, y = Crabs), size = 2, alpha = 0.4) +
  ylab("Probability of presence") +
  xlab("Temp (C)") +
  theme_classic()
```

One of the problems with using the wrong model, your standard errors become very low. Which is why the the standard error on the "correct" model here, the standard errors are larger. 

```{r}
m.best <- m.list$nb3
summary(m.best)

#deviance and degrees of freedom
dev <- summary(m.best)$deviance     # deviance
dof <- summary(m.best)$df.residual  #residual deg. of freedom

#Chi-square test (>0.05 indicates adequate fit)
1 - pchisq(dev, dof)
```

Close to .05 which is fine, but meh. Can publish but should be noted. Possible that a 0 inflated model is what you want. 