Class example using propotional Data

```{r}
set.seed(123)

#simulate 50 10-trial coin flips with pr(heads) = 0.5
heads <- rbinom(n = 50, size = 10, prob = 0.5)
heads
table(heads)
hist(heads, col = "darkseagreen1")

#simulate 10,000 10-trial coin flips with pr(heads) = 0.5
heads <- rbinom(n = 10000, size = 10, prob = 0.5)
heads
table(heads)
hist(heads, col = "darkseagreen1")
```

Take more samples, get closer to the true mean

Two types of data can be given, amount of sucesses or failures OR proportional data

```{r}
p <- 0.25 #per trial success probability
trials <- 10 #trials
samples <- 50

y <- rbinom(samples, trials, p) #binomial random variables
y.prop <- y/trials #raw proportions
y.mat <- cbind(success = y, failure = trials-y) #cbind(success, failures)

library(boot)

#proportion approach (y.prop)
prop <- glm(y.prop ~ 1, family = binomial)
summary(prop)$coefficients

inv.logit(summary(prop)$coefficients[1])

#success/failure approach (y.mat)
mat <- glm(y.mat ~ 1, family = binomial)
summary(mat)$coefficients

inv.logit(summary(mat)$coefficients[1])
```
Estimates are the same, which is good! Coefficients should be the same. average probability of success is close to our probability which is good! The standard error is different though. Trial size is explicit. 1/2 is not the same as 1000/2000. More confidence with larger trial size (therefore it's better to use trial size)

What happens when you have 3 or more possible responses?
-Choices A B C, never sometimes always

If your choices aren't ordered -> multinomial logistic regression
If your choices are ordered -> ordinal logistic regression
