Class examples of Zero-Inflated Poisson

```{r, include = TRUE}
set.seed(123)
n <- 100
pr.occ <- 0.6
lambda <- 3
occ <- rbinom(n, 1, pr.occ)   #similate occpuancy -60 ls
abund <- rpois(n, lambda*occ) #ZI counts
table(abund)/n
```

If we graphed this, there are too many zeros! we have to find the ones from binary and poisson and sepparate them.

```{r}
library(pscl)
zip <- zeroinfl(abund ~ 1 | 1, dist = "poisson")
summary(zip)[[1]]

exp(zip$coefficients$count) #mean abundance
plogis(zip$coefficients$zero) #probability of structural zeros
```

Left side is model for counts, right side is the excess zeros. Above is null model for both. Numbers match up well. 

What if we suspect zero-inflation and overdispersion?

```{r}
zinb <- zeroinfl(abund ~ 1 | 1, dist = "negbin")
summary(zinb)

exp(zinb$coefficients$count) #mean abundance
plogis(zinb$coefficients$zero) #probability of structural zeros
```

We now have a log theta term now... Theta is our dispersion factor. Negative binomial has an additional factor. Very similar to our ZIP model, how do we decide which is better? Model selection!

Comparing models: ZI models are still GLMs therefore we can compare using AIC.

```{r}
library(AICcmodavg)
aictab(list(zip, zinb), modnames = c("ZIP", "ZINB"))
```

The ZIP is better because ZINB has an extra term (theta) that is accounted for. 