
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(boot)
library(AICcmodavg)
library(pscl)
library(MuMIn)
library(ggplot2)
library(MASS)
library(stats)
options(scipen = 999)
```

```{r}
crabs <- read.table("crabs.txt", h=T)
crabs$AirTemp2 <- crabs$AirTemp^2
```

## Set list

What are the models we can use?
```{r fitlist, echo=FALSE}
# Poisson
m.null <- glm(Crabs ~ 1, data = crabs, family = "poisson")
m.p1 <- glm(Crabs ~ Site, data = crabs, family = "poisson")
m.p2 <- glm(Crabs ~ Site + AirTemp, data = crabs, family = "poisson")
m.p3 <- glm(Crabs ~ Site + AirTemp + AirTemp2, data = crabs, family = "poisson")
m.p4 <- glm(Crabs ~ AirTemp, data = crabs, family = "poisson")
m.p5 <- glm(Crabs ~ AirTemp + AirTemp2, data = crabs, family = "poisson")

# NB
m.nb1 <- glm.nb(Crabs ~ Site, data = crabs)
m.nb2 <- glm.nb(Crabs ~ Site + AirTemp, data = crabs)
m.nb3 <- glm.nb(Crabs ~ Site + AirTemp + AirTemp2, data = crabs)
m.nb4 <- glm.nb(Crabs ~ AirTemp, data = crabs)
m.nb5 <- glm.nb(Crabs ~ AirTemp + AirTemp2, data = crabs)

# ZIP nulls
m.zip1 <- zeroinfl(Crabs ~ Site | 1, data = crabs, dist = "poisson", link = "logit")
m.zip2 <- zeroinfl(Crabs ~ Site + AirTemp | 1, data = crabs, dist = "poisson", link = "logit")
m.zip3 <- zeroinfl(Crabs ~ Site + AirTemp + AirTemp2 | 1, data = crabs, dist = "poisson", link = "logit")
m.zip4 <- zeroinfl(Crabs ~ AirTemp | 1, data = crabs, dist = "poisson", link = "logit")
m.zip5 <- zeroinfl(Crabs ~ AirTemp + AirTemp2 | 1, data = crabs, dist = "poisson", link = "logit")
m.zip6 <- zeroinfl(Crabs ~ Site | 1, data = crabs, dist = "poisson", link = "logit")

m.zip7 <- zeroinfl(Crabs ~ Site | Site, data = crabs, dist = "poisson", link = "logit")
m.zip8 <- zeroinfl(Crabs ~ Site + AirTemp | Site, data = crabs, dist = "poisson", link = "logit")
m.zip9 <- zeroinfl(Crabs ~ Site + AirTemp + AirTemp2 | Site, data = crabs, dist = "poisson", link = "logit")
m.zip10 <- zeroinfl(Crabs ~ AirTemp | Site, data = crabs, dist = "poisson", link = "logit")
m.zip11 <- zeroinfl(Crabs ~ AirTemp + AirTemp2 | Site, data = crabs, dist = "poisson", link = "logit")
m.zip12 <- zeroinfl(Crabs ~ Site | Site, data = crabs, dist = "poisson", link = "logit")

m.zip13 <- zeroinfl(Crabs ~ Site | AirTemp, data = crabs, dist = "poisson", link = "logit")
m.zip14 <- zeroinfl(Crabs ~ Site + AirTemp | AirTemp,, data = crabs, dist = "poisson", link = "logit")
m.zip15 <- zeroinfl(Crabs ~ Site + AirTemp + AirTemp2 | AirTemp,, data = crabs, dist = "poisson", link = "logit")
m.zip16 <- zeroinfl(Crabs ~ AirTemp | AirTemp,, data = crabs, dist = "poisson", link = "logit")
m.zip17 <- zeroinfl(Crabs ~ AirTemp + AirTemp2 | AirTemp,, data = crabs, dist = "poisson", link = "logit")
m.zip18 <- zeroinfl(Crabs ~ Site | AirTemp,, data = crabs, dist = "poisson", link = "logit")

# ZINB
m.zinb1 <- zeroinfl(Crabs ~ Site | 1, data = crabs, dist = "negbin", link = "logit")
m.zinb2 <- zeroinfl(Crabs ~ Site + AirTemp | 1, data = crabs, dist = "negbin", link = "logit")
m.zinb3 <- zeroinfl(Crabs ~ Site + AirTemp + AirTemp2 | 1, data = crabs, dist = "negbin", link = "logit")
m.zinb4 <- zeroinfl(Crabs ~ AirTemp | 1, data = crabs, dist = "negbin", link = "logit")
m.zinb5 <- zeroinfl(Crabs ~ AirTemp + AirTemp2 | 1, data = crabs, dist = "negbin", link = "logit")
m.zinb6 <- zeroinfl(Crabs ~ Site | 1, data = crabs, dist = "negbin", link = "logit")

m.zinb7 <- zeroinfl(Crabs ~ Site | Site, data = crabs, dist = "negbin", link = "logit")
m.zinb8 <- zeroinfl(Crabs ~ Site + AirTemp | Site, data = crabs, dist = "negbin", link = "logit")
m.zinb9 <- zeroinfl(Crabs ~ Site + AirTemp + AirTemp2 | Site, data = crabs, dist = "negbin", link = "logit")
m.zinb10 <- zeroinfl(Crabs ~ AirTemp | Site, data = crabs, dist = "negbin", link = "logit")
m.zinb11 <- zeroinfl(Crabs ~ AirTemp + AirTemp2 | Site, data = crabs, dist = "negbin", link = "logit")
m.zinb12 <- zeroinfl(Crabs ~ Site | Site, data = crabs, dist = "negbin", link = "logit")

m.zinb13 <- zeroinfl(Crabs ~ Site | AirTemp, data = crabs, dist = "negbin", link = "logit")
m.zinb14 <- zeroinfl(Crabs ~ Site + AirTemp | AirTemp, data = crabs, dist = "negbin", link = "logit")
m.zinb15 <- zeroinfl(Crabs ~ Site + AirTemp + AirTemp2 | AirTemp, data = crabs, dist = "negbin", link = "logit")
m.zinb16 <- zeroinfl(Crabs ~ AirTemp | AirTemp, data = crabs, dist = "negbin", link = "logit")
m.zinb17 <- zeroinfl(Crabs ~ AirTemp + AirTemp2 | AirTemp, data = crabs, dist = "negbin", link = "logit")
m.zinb18 <- zeroinfl(Crabs ~ Site | AirTemp, data = crabs, dist = "negbin", link = "logit")

m.list <- list(
  "null" = m.null,
  "p1" = m.p1,
  "p2" = m.p2,
  "p3" = m.p3,
  "p4" = m.p4,
  "p5" = m.p5,
  "nb1" = m.nb1,
  "nb2" = m.nb2,
  "nb3" = m.nb3,
  "nb4" = m.nb4,
  "nb5" = m.nb5,
  "zip1" = m.zip1,
  "zip2" = m.zip2,
  "zip3" = m.zip3,
  "zip4" = m.zip4,
  "zip5" = m.zip5,
  "zip6" = m.zip6,
  "zip7" = m.zip7,
  "zip8" = m.zip8,
  "zip9" = m.zip9,
  "zip10" = m.zip10,
  "zip11" = m.zip11,
  "zip12" = m.zip12,
  "zip13" = m.zip13,
  "zip14" = m.zip14,
  "zip15" = m.zip15,
  "zip16" = m.zip16,
  "zip17" = m.zip17,
  "zip18" = m.zip18,
  "zinb1" = m.zinb1,
  "zinb2" = m.zinb2,
  "zinb3" = m.zinb3,
  "zinb4" = m.zinb4,
  "zinb5" = m.zinb5,
  "zinb6" = m.zinb6,
  "zinb7" = m.zinb7,
  "zinb8" = m.zinb8,
  "zinb9" = m.zinb9,
  "zinb10" = m.zinb10,
  "zinb11" = m.zinb11,
  "zinb12" = m.zinb12,
  "zinb13" = m.zinb13,
  "zinb14" = m.zinb14,
  "zinb15" = m.zinb15,
  "zinb16" = m.zinb16,
  "zinb17" = m.zinb17,
  "zinb18" = m.zinb18
)
```

Comparing through AIC table. Have to do something special to get around comparing across different models
```{r}
## AICc
aic.table <- data.frame(AIC = sapply(m.list, AIC))

aic.table$Delta.AIC <- aic.table$AIC - min(aic.table$AIC)
aic.table <- aic.table[order(aic.table$AIC),]
aic.table
```

Best model is zinb6 now


## Visualizing the data with predict function
```{r predict}
m.best <- m.list$zinb14
summary(m.best)

crabs.newdata <- expand.grid(seq(min(crabs$AirTemp),max(crabs$AirTemp),0.1), levels(as.factor(crabs$Site)))
colnames(crabs.newdata) <- c("AirTemp","Site")
crabs.newdata$AirTemp2 <- crabs.newdata$AirTemp^2

crabs.newdata$pred.counts <- predict(m.best,crabs.newdata,type="response")
crabs.newdata$pred.zeros <- predict(m.best,crabs.newdata,type="zero")

ggplot(crabs.newdata, aes(AirTemp, pred.counts, color = Site)) +
  geom_line() +
  labs(title = "Predicted counts", y = "Predicted number of crabs") +
  theme_classic()

ggplot(crabs.newdata, aes(AirTemp, pred.zeros, color = Site)) +
  geom_line() +
  labs(title = "Predicted zeros", y = "Predicted proportion of zeros") +
  theme_classic()
```

To read the coefficient estimates! When you get the summary back, you get the values back in a raw form. You have to translate the values back to get it into crab counts (we use the predict function for this). From there, coefficients are as usually. (constrats). Looking at the count model, estimates from the sites are crab counts when temperature = 0 (intercept) and the airtemp estimates are slope values. 